## Functions and Conditionals

### Conditionals

The basic syntax for an if statement goes like this:

```{r}
if (logical statement) {
  do something
} else {
  do something else
}
```

The logical statements should boil down to a boolean; for example:

```{r}

```

We can also add else if statements in the middle if we need more options.

```{r}

```

Neat! We can also nest conditionals just like we do in Python.

```{r}

    
```

Neat! Now, let's look at another core part of R programming - Loops!

### Loops

```{r}

```

We can also do nifty stuff with our vectors, such as apply changes to specific indices of our data.

```{r}

```

We can loop over more than just lists:

```{r}

```

Like Python, we have while loops in R:

```{r}

```

## Functions

When should we use functions?

```{r}
our_function <- function(optional_params){
  # do some stuff
}
```

An example:

```{r}

```

This is all cool.. but it didn't do anything. We need to call the function!

```{r}

```

Notice that in the background here, we are looping over each item in the array, but we never write a for loop! Keep this in mind, since when you pass in a large data set, the changes made by functions aren't quite immediate - the computer does have to go over each item in your data set, and change them one at a time.

A shortcut of sorts is to use the `apply()` function, which operates very similarly to Python's `map()`. Note that `apply` syntax goes as follows: `apply(X, MARGIN, FUN)` where X is the data set to use, MARGIN is specification for rows or columns, and FUN is the function to apply.

```{r}

```

It is common to use `mean`, `median`, `sum`, `min`, or `max` with `apply`, but we can also use our user-defined functions as well. Note that apply() only works with matrices.

```{r}

```

### Homework

1.  Do 5 CodeWars problems in R. Share your solutions here:

    ```{r}
et_chance <- function(n, x, a) {
  total_outcomes <- choose(n, a)

  favorable_outcomes <- choose(n - x, a)

  P <- favorable_outcomes / total_outcomes

  P <- round(P, 2)

  return(P)
}
    array_diff <- function(a, b) {
  # Select the elements of 'a' that are not in 'b'
  result <- a[!a %in% b]
  
  return(result)
}
disemvowel <- function(string) {
  result <- gsub(pattern = "[aeiouAEIOU]", replacement = "", x = string)
  
  return(result)
}
row_sum_odd_numbers <- function(n) {
  result <- n^3
  
  return(result)
}
nbDig <- function(n, d) {
  # Initialize the count to 0
  count <- 0
  
  # Loop over each number from 0 to 'n'
  for (k in 0:n) {
    # Square the number
    square <- k^2
    
    # Convert the squared number to a string
    square_string <- as.character(square)
    
    # Count the occurrences of the digit 'd' in the string
    count <- count + stringr::str_count(square_string, pattern = as.character(d))
  }
  
  return(count)
}


    ```

2.  Create a function that checks an input of type `int` and returns fizz if the number is a multiple of 3, buzz if the number is a multiple of 5, and fizzbuzz if the number is a multiple of 3 and 5.

    ```{r}
fizzbuzz <- function(num) {
  if (num %% 3 == 0 & num %% 5 == 0) {
    return("fizzbuzz")
  } else if (num %% 3 == 0) {
    return("fizz")
  } else if (num %% 5 == 0) {
    return("buzz")
  } else {
    return(num)
  }
}

result <- fizzbuzz(15)
print(result) 

    ```

3.  **Mario Challenge:** In the classic Super Mario games, Mario ends each level by ascending a flight of stairs and leaping off the top to arrive at the castle at the end. (See [here](https://youtu.be/-avspZlbOWU?t=55) for details). We can imagine that each of the bricks of the stairs is a character X. Create a function that prints out a set of stairs. For example:

    If the function is called marioStairs() and we pass in the value 4, we would expect to see stairs with a length of 4 and a height of 4:

    X

    XX

    XXX

    XXXX

    ```{r}
    # You may feel more comfortable figuring this out in Python first, then translating into R later.
    marioStairs <- function(n) {
  for (i in 1:n) {
    cat(rep("X", i), "\n")
  }
}

# Example usage:
marioStairs(4)

    ```
