title: "R Notebook"
output: html_notebook
---

# R: Data Structures

## Vectors


```{r}
# Create a vector
my_vector <- c(1, 2, 3, 4, 5)
print(my_vector)

# Get the type of the vector
print(typeof(my_vector))

Getting a Vector’s Length
You can get the length of a vector using the length() function

# Get the length of the vector
print(length(my_vector))




# Note this returns "John", the first element in our Vector
print(my_vector[1])

# Generate a sequence from 5 to 15
seq1 <- 5:15
print(seq1)

# Generate a sequence with decimals
seq2 <- seq(1, 2, by = 0.1)
print(seq2)


# Step from 5 to 15 by increments of .3
seq3 <- seq(5, 15, by = 0.3)
print(seq3)


```

### Calculations between two matrices

# Get the element in the 2nd row and 4th column
print(my_matrix[2, 4])

You can also get entire rows and columns.

# Get the second row
print(my_matrix[2, ])

# Get the third column
print(my_matrix[, 3])

Calculations between two matrices
# Create two matrices
matrix1 <- matrix(1:9, nrow = 3, ncol = 3)
matrix2 <- matrix(10:18, nrow = 3, ncol = 3)

Addition
# Add the matrices
result_add <- matrix1 + matrix2
print(result_add)

Subtraction
# Subtract the matrices
result_sub <- matrix1 - matrix2
print(result_sub)

Multiplication
# Multiply the matrices
result_mul <- matrix1 * matrix2
print(result_mul)

Division
# Divide the matrices
result_div <- matrix1 / matrix2
print(result_div)
## Arrays

Arrays can store data in more than two dimensions.They are n-dimensional. If we create an array with the dimensions of (5,6,6), it will create 6 matrices with 5 rows and 6 columns. Arrays are created with the `array ()` function. The first argument is the vector of data to display, the second is the desired dimensions for your array.

```{r}

```

### Naming Array columns and rows

You can name columns, rows, and matrices using the optional dimnames parameter

```{r}
# Name the columns, rows, and matrices
dimnames(my_array) <- list(c("Row1", "Row2", "Row3"), c("Col1", "Col2", "Col3"), c("Matrix1", "Matrix2"))
print(my_array)

```

### Accessing Array Elements

# Get the element in the second row, third column, first matrix
print(my_array[2, 3, 1])

# Get the second row of all of the matrices
print(my_array[2, , ])

# Get the second column of all of the matrices
print(my_array[, 2, ])

# Get the second matrix
print(my_array[, , 2])

```

# A factor of eye colors
eye_colors <- factor(c("Blue", "Green", "Blue", "Brown", "Green", "Brown", "Blue"))
print(eye_colors)

You can also use the optional levels parameter to specify the order in which the levels should be displayed.

# Specify the order of the levels
eye_colors <- factor(eye_colors, levels = c("Blue", "Green", "Brown"))
print(eye_colors)
```

## Homework

1.  Create an array of two 3x3 matrices from the given vectors.

```{r}
homework_array <- array(1:18, dim = c(3, 3, 2))
print(homework_array)

```

2.  In the above matrix that you created, print the value 5 by indexing in to the matrix.
# Print the value 5
print(homework_array[2, 2, 1])

3.  Print the third row of the first matrix.
# Print the third row of the first matrix
print(homework_array[3, , 1])

4.  Print the second column of both matrices.

    ```{r}
# Print the second column of both matrices
print(homework_array[, 2, ])

    ```
